---
title: "Chicago Bike Rack Kernel Density"
author: "Aisha Syed"
date: '2022-11-11'
output: html_document
---

## Intoduction

The following notebook contains code to create a kernel density map of bike racks in Chicago, Illinois.

```{r, message=FALSE}
#set up environment
setwd("~/GEOG224/Assignment4_Kringing")
library(sf)
library(tmap)
library(raster) # Needed for grid and kernel density surface
library(adehabitatHR) # Needed for kernel density surface
```

## Load and Clean Data
Now, I will load in data. The data was acquired from the following sources: <br>
Chicago neighborhood boundaries (2018): Chicago Data Portal / City of Chicago <br>
Bike rack locations (2012): Chicago Data Portal / City of Chicago

```{r}
chi = st_read("~/GEOG224/Assignment4_Kringing/data/Boundaries - Neighborhoods.geojson")
bikeracks = read.csv("data/Bike_Racks.csv")

#inspect chicago neighborhoods
st_crs(chi) #4326
tm_shape(chi) + tm_polygons()

#change the crs into a projection that preserves area
chi = st_transform(chi, 26971)
tm_shape(chi) + tm_polygons()

#inspect bike rack data
str(bikeracks) #5164 observations (nonspatial)
```

It seems that the bike rack data already has numeric latitude and longitude data. We just have to convert the csv into spatial data.

```{r}
bikeracks_sf = st_as_sf(bikeracks, coords = c("Longitude", "Latitude"), crs = 4326) |>
  st_transform(26971)

str(bikeracks_sf) #5164 observations

#see if there are any NA values
subset(bikeracks_sf, is.na(bikeracks_sf[,"geometry"])) #There are no row with empty geometry! Yay!
```

## Plot Point Data
Below, I will create a map using tmap that displays the bike racks in Chicago.
```{r}
tm_shape(chi) + tm_polygons(col = "white")+
  tm_shape(bikeracks_sf) + tm_dots(col = "pink") +
  tm_layout(main.title = "Chicago Bike Racks (2012)", main.title.size = 1, bg.color = "grey85") + tm_scale_bar(position=c("left", "bottom"))
```

## Create a Kernel Density Map
Below, I created a kernel density map in Chicago to show the density of bike racks. I used href as the h parameter with a Epanechnikov kernel.

```{r}
#convert sf to sp
bikeracks_sp = sf:::as_Spatial(bikeracks_sf)
chi_sp = sf:::as_Spatial(chi)

#create KDE with different h parameters
kde_output_href = kernelUD(bikeracks_sp, h="href", grid = 1000)
kde_output_lscv = kernelUD(bikeracks_sp, h="LSCV", grid = 1000)

plot(kde_output_href)
plot(kde_output_lscv)
```

I like when the h parameter is href, because it provides a clearer pattern opposed to when the h parameter is LSCV. From now on, I will utilize the result when the h parameter is href. Now, I will see which kernel I would like to use.

```{r, warning=FALSE}
#create KDE with different kernels
kde_output_biv = kernelUD(bikeracks_sp, h="href", kern = "bivnorm", grid = 1000)
kde_output_epa = kernelUD(bikeracks_sp, h="href", kern = "epa", grid = 1000)

plot(kde_output_biv)
plot(kde_output_epa)
```
I like using Epanechnikov kernel when the kern parameter = "epa" because it shows a distinct pattern on the north side of Chicago.

```{r}
#make the output as a raster
kde = raster(kde_output_epa)
bounding_box = bbox(chi_sp)

masked_kde = mask(kde, chi)

tm_shape(masked_kde, bbox = bounding_box) + tm_raster("ud", style = "quantile", n = 100, legend.show = F, palette = "RdPu") +
  tm_shape(chi_sp) + tm_borders(alpha=.3, col = "white") +
  tm_layout(main.title = "Chicago Bike Rack Density (2012)",  main.title.size = 1, frame = FALSE)+ tm_scale_bar(position=c("left", "bottom"))

```

## Create a Contour Map

Below, I will create contours that reflect the density of bike racks in Chicago.

```{r}
#create contour ranges
range75 <- getverticeshr(kde_output_epa, percent = 75)
range50 <- getverticeshr(kde_output_epa, percent = 50)
range25 <- getverticeshr(kde_output_epa, percent = 25)

#make map
tm_shape(chi) + tm_fill(col = "#f0f0f0") + tm_borders(alpha=.8, col = "white") +
  tm_shape(range75) + tm_borders(alpha=.7, col = "hotpink1", lwd = 2) + tm_fill(alpha=.1, col = "hotpink1") +
  tm_shape(range50) + tm_borders(alpha=.7, col = "hotpink3", lwd = 2) + tm_fill(alpha=.1, col = "hotpink3") +
  tm_shape(range25) + tm_borders(alpha=.7, col = "hotpink4", lwd = 2) + tm_fill(alpha=.1, col = "hotpink4") +
  tm_layout(main.title = "Chicago Bike Rack Density (2012)",  main.title.size = 1, bg.color = "grey85") + tm_scale_bar(position=c("left", "bottom"))
```




